---
title: "Ubiquitous genes"
author: "Luke Hayden"
date: "March 2016"
output: html_document
---

#Importing our datasets
File:
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
setwd("~/Documents/expsummary")

#source("https://bioconductor.org/biocLite.R")
#biocLite("EBSeq")
#biocLite("DESeq2")

library(ggplot2)
library(DESeq2)
library(rmarkdown)
library(pracma)
library(seqinr)
library(ape)
library(reshape2)
library(scales)


# collect the names of the files with count data. there are 16 files, that correspond to 8 samples sequenced twice 
lf=Sys.glob("*.count")

# read these files sequentially and return the column with counts
lht=sapply(lf,function(x){
ht=read.table(x,h=T);return(ht[,2])})
lht=data.frame(lht)

# make a dataframe with 16 columns and the row names with gene ids
ht=read.table(lf[[1]],h=T);
row.names(lht)=ht[,1]

# identify the columns with the number of lane and the sample id
id=rep(1:8,2)
lane=c(rep("L03",4),rep("L04",4),rep("L05",4),rep("L06",4))
seq=c(rep("seq1",8),rep("seq2",8))
type=rep(c("O1","O2","Y1","Y2","RO1","RO2","RY1","RY2"),2)

# count the number of reads per sample
nb=apply(lht,2,sum)

# retain only the lines corresponding to genes
lht1=lht[1:28666,]
nb1=apply(lht1,2,sum)


```
The transcriptomic dataset used here is the 8-sample aging/regeneration one. 
Samples are either:

Young or old, regenerated or not, replicate 1 or 2. 


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

lht2=lht[28667:28671,]
lht3=as.matrix(rbind(lht2[c(1,3,5),],nb1))
ht=lht1[,1:8]+lht1[,9:16]
colnames(ht) <- c("O1", "O2", "Y1", "Y2", "OR1", "OR2", "YR1", "YR2")
```
We perform the following tests of differential expression (aging dataset):

-DE old vs young

-DE regenerated vs not

-DE between the two replicates of each of the four conditions
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
#Age
colData <- data.frame(age=c("O","O","Y","Y","O","O","Y","Y"),type=c(rep("NR",4),rep("R",4)),rep=c(rep(c("rep1","rep2"),4)))

dds <- DESeqDataSetFromMatrix(countData = ht,
colData = colData,
design = ~ age)

#DE age
ddsA <- DESeqDataSetFromMatrix(countData = ht[,colData$type=="NR"],
colData = colData[colData$type=="NR",],
design = ~ age)
ddsA <- DESeq(ddsA)
resA05 <- results(ddsA, alpha=0.05)

#DE regen
ddsR <- DESeqDataSetFromMatrix(countData = ht[,colData$age=="Y"],
colData = colData[colData$age=="Y",],
design = ~ type)
ddsR <- DESeq(ddsR)
resR05 <- results(ddsR, alpha=0.05)



#DE OR1 vs 2
ddsOR <- DESeqDataSetFromMatrix(countData = ht[,colData$type=="R"&colData$age=="O"],
colData = colData[colData$type=="R"&colData$age=="O",],
design = ~ rep)
ddsOR <- DESeq(ddsOR,fitType='local')
resOR01 <- results(ddsOR, alpha=0.1)
nameResOR01=row.names(resOR01[!is.na(resOR01$padj)&resOR01$padj<0.1,])

#DE YR 1 vs 2
ddsYR <- DESeqDataSetFromMatrix(countData = ht[,colData$type=="R"&colData$age=="Y"],
colData = colData[colData$type=="R"&colData$age=="Y",],
design = ~ rep)
ddsYR <- DESeq(ddsYR,fitType='local')
resYR01 <- results(ddsYR, alpha=0.1)
nameResYR01=row.names(resYR01[!is.na(resYR01$padj)&resYR01$padj<0.1,])


#DE ONR 1 vs 2
ddsONR <- DESeqDataSetFromMatrix(countData = ht[,colData$type=="NR"&colData$age=="O"],
colData = colData[colData$type=="NR"&colData$age=="O",],
design = ~ rep)
ddsONR <- DESeq(ddsONR,fitType='local')
resONR01 <- results(ddsONR, alpha=0.1)
nameResONR01=row.names(resONR01[!is.na(resONR01$padj)&resONR01$padj<0.1,])

#DE YNR 1 vs 2
ddsYNR <- DESeqDataSetFromMatrix(countData = ht[,colData$type=="NR"&colData$age=="Y"],
colData = colData[colData$type=="NR"&colData$age=="Y",],
design = ~ rep)
ddsYNR <- DESeq(ddsYNR,fitType='local')
resYNR01 <- results(ddsYNR, alpha=0.1)
nameResYNR01=row.names(resYNR01[!is.na(resYNR01$padj)&resYNR01$padj<0.1,])


```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
# collect the names of the files with .tsv data. there are 4 files
lf=Sys.glob("*.tsv")


#read these in
reads <- sapply(lf, function(x){
r=read.table(x, skip=1);return(r$V4)})
reads <- data.frame(reads)


desc <- read.table("NKS-01_6_S1_abundance.tsv", h=T)
row.names(reads) <- desc$target_id

colnames(reads) <- c("Before_Amputation", "Regenerated", "Amputated+3d", "Amputated+6d")

rreads <- round(reads)

sums <- colSums(reads)
```
We then convert the counts into RPKMs
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
#Calculating RPKMs
htRPKMs <- ht
NbyL <- as.data.frame(desc$length %o% colSums(ht))
htRPKMs <- htRPKMs/NbyL
htRPKMs <- htRPKMs*1000000000

mr <- rowMeans(htRPKMs)
quantile(mr, 0.95)

thtr <- subset(htRPKMs, mr> quantile(mr, 0.95))



readsRPKMs <- reads
NbyL <- as.data.frame(desc$length %o% sums)
readsRPKMs <- readsRPKMs/NbyL
readsRPKMs <- readsRPKMs*1000000000

rmr <- rowMeans(readsRPKMs)

```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}



el <- data.frame(mr, isho =mr> quantile(mr, 0.9), rmr, ishr = rmr> quantile(rmr, 0.9))



#ggplot(el, aes(x=mr, y=rmr, colour=isho, fill=ishr)) +geom_point(shape=21) +scale_y_log10() +scale_x_log10() +ylab("Mean Expression in Regeneration Dataset (RPKMs)") +xlab("Mean Expression in Aging Dataset (RPKMs)") + scale_colour_manual(values=c("orangered", "forest green"), breaks=c(TRUE, FALSE), labels=c(">90th centile", "<90th centile"), name="Meaning Aging Dataset Expression") +scale_fill_manual(values=c("black", "white"),name="Mean Regenerating Dataset Expression", labels=c(">90th centile", "<90th centile"))



subset(el, el$isho==TRUE & el$ishr ==TRUE)

```
#Imposing cutoffs


Let
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
cutoff <- 0.8

isc<- htRPKMs[,1]> quantile(htRPKMs[,1], cutoff)

for (i in 1:length(colnames(htRPKMs))){
  isc <- isc & (htRPKMs[,i]> quantile(htRPKMs[,i], cutoff))
}


for (i in 1:length(colnames(readsRPKMs))){
  isc <- isc & (readsRPKMs[,i]> quantile(readsRPKMs[,i], cutoff))
}
pc <- toString((1-cutoff)*100)

mhtRPKMs <- melt(cbind(htRPKMs,readsRPKMs, isc))
ggplot(mhtRPKMs, aes(y=value, x=variable, colour=isc)) +geom_point(alpha=0.15, shape=95, size=8) +scale_y_log10() +scale_colour_manual(values=c("cornflower blue", "red3"),breaks=c(TRUE, FALSE), labels=c(">80th centile all samples", "<80th centile"), name="Gene cut-off") +ylab("Log Expression level (RPKMs)") +ggtitle(paste(sum(isc), "genes in top", toString((1-cutoff)*100), "% expression for all samples" ))  +xlab("Transcriptomic Sample") + theme(axis.text.x = element_text(angle = 60, hjust = 1))

#sum(isc)

```


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
lfc <- data.frame(resA05$log2FoldChange, resR05$log2FoldChange, resONR01$log2FoldChange, resOR01$log2FoldChange, resYNR01$log2FoldChange, resYR01$log2FoldChange)
rownames(lfc) <- rownames(resA05)
lfc$means <- rowMeans(abs(lfc))
lfc <- subset(lfc, rownames(lfc) %in% rownames(thtr))

mg <- data.frame(mnexp = rowMeans(thtr), var=lfc$means )
mg$isnotvar <- rownames(mg) %in% rownames(mg[order(mg$var)[1:50],])



#plot(lfc$means, rowMeans(thtr), ylim=c(0,2000))

#ggplot(mg, aes(y=mnexp, x=var, colour=isnotvar)) +geom_point() + scale_y_log10(breaks =trans_breaks("log10", function(x) 10^x)) + ggtitle("Top 5% most expressed genes") +xlab("Variability (Mean log2foldchange between conditions & replicates)") +ylab("Mean Expression (RPKMs)") +scale_colour_manual(values=c("darkblue", "forest green"), labels=c("Other", "Top 50"), name="Marker Unvariability") +theme_bw() 

```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
sl <- rownames(subset(mg, mg$isnotvar ==TRUE ))
fastaprot=read.fasta(file="phaw.3.0.genes.prot.fa")


cfastaprot <- subset(fastaprot, names(fastaprot) %in%sl)

cfastaprot <- cfastaprot[order(names(cfastaprot))]

write.fasta(cfastaprot,names=names(cfastaprot),file=paste("ubiqgenesprot.fa", sep=""))



```


#BLAST our genes (from Parhyale) against a set of arthropod protein sets
Output these results to:
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

#system("makeblastdb -in arthropodSet.fasta -dbtype 'prot' -out arth1")

#system("blastp -db arth1 -query ubiqgenesprot.fa -outfmt '6 qseqid sseqid pident length evalue bitscore stitle' > vsarthblast.txt" )

vsarthBlast <- read.table("vsarthblast.txt", sep="\t")

colnames(vsarthBlast) <- c("Query", "Target", "Percent_ID", "Ali_len", "E_val", "bitscore", "Hit_name_long" )

write.csv(vsarthBlast, file =paste("ubiqvsarthBlastResults", fileID, ".csv"))

```



```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}


cgTopBlast <- subset(vsarthBlast, !(duplicated(vsarthBlast$Query)))
print(cgTopBlast$Hit_name_long)

```



